---
title: "ReadMe - run_analysis.R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting and Cleaning Data Course Project

This explains the steps in the script "run_analysis.R"

##### Fetch metadata
```{r eval=FALSE}
activity_labels <- 
    read.csv("data/assignment/activity_labels.txt", 
             header = FALSE, sep = "", col.names = c("activityid", "activity"))
features <- 
    read.csv("data/assignment/features.txt", 
             header = FALSE, sep = "")[2] ## Ignore first column
```

##### Fetch training data
```{r eval=FALSE}
subject_train <- read.csv("data/assignment/train/subject_train.txt", 
                          header = FALSE, col.names = "subjectid")

y_train <- read.csv("data/assignment/train/y_train.txt", 
                    header = FALSE, col.names = "activityid")

X_train <- read.csv("data/assignment/train/X_train.txt", 
                    header = FALSE, sep = "")
```
##### Fetch test data
```{r eval=FALSE}
subject_test <- read.csv("data/assignment/test/subject_test.txt", 
                         header = FALSE, col.names = "subjectid")

y_test <- read.csv("data/assignment/test/y_test.txt", 
                   header = FALSE, col.names = "activityid")

X_test <- read.csv("data/assignment/test/X_test.txt", 
                   header = FALSE, sep = "")
```
##### Merge the training and the test sets to create one data set
```{r eval=FALSE}
subjects <- rbind(subject_train, subject_test)
activities <- rbind(y_train, y_test)
dataset0 <- rbind(X_train, X_test)
```

##### Use descriptive column names
```{r eval=FALSE}
names(dataset0) <- features[,1]
```

##### Extract the measurements on the mean and standard deviation for each measurement
All columns with names that contain the strings "mean" or "std"
```{r eval=FALSE}
statsColumns <- grepl("mean|std", names(dataset0), ignore.case = TRUE)
dataset <- dataset0[statsColumns]
```

##### Attach subjects & activities to data
```{r eval=FALSE}
dataset <- cbind(activities, subjects, dataset)
```

##### Use descriptive activity names to name the activities in the data set
Merge() will re-orderthe data, therefore this step must be performed after activities are merged with data
```{r eval=FALSE}
dataset <- merge(dataset, activity_labels)
```
  
##### Remove column 'activityid', move column 'activity' to the front
```{r eval=FALSE}
dataset$activityid <- dataset$activity
dataset$activity <- NULL
names(dataset)[1] <- "activity"
```

##### Appropriately labels the data set with descriptive variable names


##### Create a second, independent tidy data set with the average of each variable for each activity and each subject.
```{r eval=FALSE}
dataset2 <- dataset %>% 
  group_by(subjectid, activity) %>%
  summarise_all(funs(mean))
```

##### Display the new dataset
```{r}
print(dataset2)
```
